---
title: "GPQR"
output: rmarkdown::html_vignette
vignette: Group Penalized Quantile Regression in R A VIGNETTE
  %\VignetteIndexEntry{GPQR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
knitr::opts_knit$set(global.par = TRUE)
knitr::opts_chunk$set(fig.width = 7, fig.height = 4)
```

# Installation

Like many other R packages, the simplest way to obtain GPQR is to install it  from github. Type the following command in R console:

```{r setup}
library(devtools)
#devtools::install_github("https://github.com/ouhourane/GPQR.git")
```

In this vignette, we demonstrate how to use the GSQR() and cv.GSQR() functions in the GPQR
package to fit the regularization path of parametric quantile regression with grouped penalties. 
This includes group selection methods such as group Lasso, group Mcp, and group Scad. 
The others function: predict(), coef(), cv.predict, cv.coef are similar to those in glmnet package. 

Let’s look at a simple example:


# Bardet dataset

Gene expression data (20 genes for 120 samples) from the microarray experiments 
of mammalian eye tissue samples of Scheetz et al. (2006).
This data set contains 120 samples with 100 predictors 
(expanded from 20 genes using 5 basis B-splines, as described in Yang, Y. and Zou, H. (2012)).

```{r bardet}
library(GPQR)
#load bardet dataset
data(bardet)
group <- rep(1:20,each=5)
#run GPQR for group Lasso penaly, taux = 0.5 and  with the penalty group Lasso 
#and the first pseudo-quantile approximation loss function 
fit <- GPQR(x=bardet$x,y=bardet$y,group=group,method="GLasso",check="f1",taux=0.5)
coef(fit)
```

"fit" is an object of class GPQR that contains all the relevant information of the fitted the QR for further use.
To extract the extract the information from "fit", various functions are provided for the
this object such as plot, print, coef and predict that enable us to run easily those tasks.



```{r illustration}
library("MASS")
library("grpreg")
xlm=c(-2,4)
Ng=11
group=c(1,1,1,1,2,2,2,2,3,3,3,3,4:Ng)
n=100
p=length(group)
X=NULL
sig=3
betac=c(rep(3,4),rep(2,4),rep(-1,4),rep(0,p-12))
cc1=c(1,2,3,4);cc2=c(5,6,7,8);cc3=c(9,10,11,12);cc4=c(13:(p-1));
MuVec<-rep(0,p);v<-rep(1,p);SigmaMat<-diag(v)
for(i in 1:p){for(j in 1:p){SigmaMat[i,j]<- 0.5^abs(i-j);SigmaMat[i,i]<- 1}}
set.seed(123456789) 
x<-mvrnorm(2*n,MuVec,SigmaMat,tol = 1e-6, empirical = FALSE)
xx = NULL
for (h in 1:3) for (k in 1:4) xx=cbind(xx,x[,h]+rnorm(2*n,0,0.1))
xx = cbind(xx,x[,4:Ng])
Xtr=xx[1:n,]
Ytr<-Xtr%*%betac+pnorm(Xtr[,p])*rnorm(n,0,sig)##rnorm(n,0,sig)
dd=20;ddd=15
taux = 0.95

getmin <-function(lambda, cvm, cvsd) {
    cvmin <- min(cvm)
    idmin <- cvm <= cvmin
    lambda.min <- max(lambda[idmin])
    idmin <- match(lambda.min, lambda)
    semin <- (cvm + cvsd)[idmin]
    idmin <- cvm <= semin
    lambda.1se <- max(lambda[idmin])
    list(lambda.min = lambda.min, lambda.1se = lambda.1se)
}
GPQR_illustration <-function(penalty, taux){
  cv <- cv.GPQR(x=Xtr,y=Ytr,group=group,method=penalty,check="f1",taux=taux)
fittGL=t(cv$finalfit$beta)
seqLambda=cv$lambda
seqLambda=cv$lambda
ll = getmin(cv$lambda, cv$cv, cv$cv.error)
l = which(cv$lambda == ll$lambda.min)
main_lab = paste("Q-",penalty," ",expression(tau)," = ", taux)
matplot(seqLambda,fittGL[,cc1], type = "l",col = 3,lty = 1,ylim=xlm,lwd=1,
        ylab="Coefficients",main=main_lab, xlab = expression(lambda))
matlines(seqLambda,fittGL[,cc2], type = "l",col = 2,lty = 1,lwd=1)
matlines(seqLambda,fittGL[,cc3], type = "l",col = 4,lty = 1,lwd=1)
matlines(seqLambda,fittGL[,cc4], type = "l",col = 1,lty = 1,lwd=1)
matlines(seqLambda,fittGL[,20], type = "l",col = 6,lty = 1,lwd=1)
abline(v=seqLambda[l],col=1,lty = 2)
text(0.02,0.75,expression("G"[11]),col=6)
text(0.02,3.2,expression("G"[1]),col=3)
text(0.02,2.2,expression("G"[2]),col=2)
text(0.02,-1.4,expression("G"[3]),col=4)
}
```

```{r GPQR_tau_0.95}
par(mfrow = c(1,3))
GPQR_illustration("GLasso", 0.95)
GPQR_illustration("GMcp", 0.95)
GPQR_illustration("GScad", 0.95)
```
```{r GPQR_tau_0.50}
par(mfrow = c(1,3))
GPQR_illustration("GLasso", 0.50)
GPQR_illustration("GMcp", 0.50)
GPQR_illustration("GScad", 0.50)
```

```{r grpreg}
grpreg_illustration <-function(penalty){
cv <- cv.grpreg(Xtr, Ytr, group, penalty=penalty)
fittL=t(cv$fit$beta)[,-1]
seqLambdaL=cv$lambda
l=cv$min
matplot(seqLambdaL,fittL[,cc1], type = "l",col = 3,lty = 1,ylim=xlm,lwd=1,
        ylab="Coefficients",main=paste("LS-",penalty) ,xlab = expression(lambda))
matlines(seqLambdaL,fittL[,cc2], type = "l",col = 2,lty = 1,lwd=1)
matlines(seqLambdaL,fittL[,cc3], type = "l",col = 4,lty = 1,lwd=1)
matlines(seqLambdaL,fittL[,cc4], type = "l",col = 1,lty = 1,lwd=1)
matlines(seqLambdaL,fittL[,20], type = "l",col = 6,lty = 1,lwd=1)
###abline(v=seqLambdaL[dd],col = 1,lty = 3)
##abline(v=seqLambdaL[dd+ddd],col = 2,lty = 3)
abline(v=seqLambdaL[l],col=1,lty = 2)
text(0.4,0.75,expression("G"[11]),col=6)
text(0.4,3.3,expression("G"[1]),col=3)
text(0.4,2.3,expression("G"[2]),col=2)
text(0.4,-1.5,expression("G"[3]),col=4)
}
xlm=c(-3,7)
par(mfrow = c(1,3))
grpreg_illustration("grLasso")
grpreg_illustration("grMCP")
grpreg_illustration("grSCAD")
```

# real data set

```{r import_data}
#save.image(file = "PCEV.RData")
#load("PCEV.RData")
load(url("https://github.com/ouhourane/GPQR/raw/main/data/PCEV.RData"))
ylab = expression(beta~'value')
par(mfrow = c(2, 3))
set.seed(123)
################ package grpreg
Trace_grpreg <- function (penalty){
  cvfit = cv.grpreg(Xc, pheno, group, penalty=penalty)
  coefGM=predict(cvfit, Xc, type="coefficients")[-1]
  matplot(positionC/1e6,coefGM,col=4, type = "p",pch=19,cex=1,ylab=ylab,xlab="Location",main=paste("LS-",penalty),ylim=c(-0.008,0.006))
  abline(v=11.235,lty=2) 
  abline(v=11.385,lty=2)
}
## tracer
Trace_grpreg("grLasso")
Trace_grpreg("grMCP")
Trace_grpreg("grSCAD")

################ package GPER
Trace_GPQR <- function (penalty,taux){
  cvGS=cv.GPQR(Xc, pheno, group, Kfold = 5, taux = taux, check ="f1",method =penalty,plot.it=F)
  l=min(which(cvGS$cv==min(cvGS$cv)))
  nbeta <- c(cvGS$finalfit$b0[l], cvGS$finalfit$beta[,l])
  matplot(positionC/1e6,nbeta[-1],col=4, type = "p",pch=19,cex=1,ylab=ylab,xlab="Location",
          main= paste("GPQR -",penalty," tau=",taux),ylim=c(-0.008,0.006))
  abline(v=11.235,lty=2) 
  abline(v=11.385,lty=2)
}
## tracer
Trace_GPQR("GLasso",0.5)
Trace_GPQR("GMcp",0.5)
Trace_GPQR("GScad",0.5)
```





